---
title: "Performance Metrics of Chemical Processing: A Regression Analysis of Belle Ayr Data"
author: "Joshua Berthiaume"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

# 1. Loading the Data

First, we load the data from a CSV file.

```{r}
dat <- read.csv("data.csv")
```

# 2. Building the Multiple Regression Model

Next, we create a multiple regression model to relate CO2 (y) to total solvent (x6) and hydrogen consumption (x7).

```{r}
model <- lm(y ~ x6 + x7, data = dat)
summary(model)
```

Equation of the fitted line:
Y^ = 2.526+0.019(x6)+2.186(x7)

# 3. Testing the Significance of the Regression Model

We perform an ANOVA to test the overall significance of the regression model.


```{r}
anova(model)
```

## 3.1 F-Test for Overall Significance

We calculate the p-value and rejection region for the F-test.

```{r}
# F-statistic p-value
p_value <- 1 - pf(27.95, 2, 24)
print(p_value)

# Rejection region
alpha <- 0.05
rejection_region <- qf(1 - alpha, 2, 24)
print(rejection_region)
```

Interpretation of ANOVA Results:

    Hypotheses:
        Null Hypothesis (H₀): Both coefficients β₆ and β₇ are zero.
        Alternative Hypothesis (Hₐ): At least one of the coefficients β₆ or β₇ is not zero.

    F-Test Statistic:
        The F-statistic is 27.95, which is much greater than the critical value of 3.40.
        We reject the null hypothesis, indicating that at least one of the predictors is significant.

    Model Fit:
        R2=0.6996R2=0.6996
        Adjusted R2=0.6746R2=0.6746

# 4. Testing the Contribution of Individual Predictors

We now test the significance of each predictor (x6 and x7) using t-tests.

## 4.1 Hypothesis Tests for Coefficients

We first calculate the critical value for the t-tests.

```{r}
n <- nrow(dat)
k <- 2 + 1
t_critical <- qt(1 - 0.05 / 2, n - k)
print(t_critical)

```
### 4.1.1 For β₆ (Total Solvent)

    Hypothesis:
        Null Hypothesis (H₀): β₆ = 0
        Alternative Hypothesis (Hₐ): β₆ ≠ 0

    Test Statistic:
        The test statistic is 6.74.

    Conclusion:
        Since 6.74 > ±2.064, we reject the null hypothesis. x6 is significantly related to y.

### 4.1.2 For β₇ (Hydrogen Consumption)

    Hypothesis:
        Null Hypothesis (H₀): β₇ = 0
        Alternative Hypothesis (Hₐ): β₇ ≠ 0

    Test Statistic:
        The test statistic is 2.247.

    Conclusion:
        Since 2.247 > ±2.064, we reject the null hypothesis. x7 is significantly related to y.

# 5. Confidence Intervals
## 5.1 95% Confidence Interval for β₆

We construct the 95% confidence interval for β₆.
```{r}
confint(model, "x6")
```

Conclusion: The 95% confidence interval for β₆ is (0.0129, 0.0242).
## 5.2 90% Confidence Interval for Mean CO2

We calculate the 90% confidence interval for the mean CO2 when x6 = 3200 and x7 = 7.2.

```{r}
predict(model,interval="confidence",newdata = data.frame(x6=3200,x7=7.2),level=.9)
```
Conclusion: The 90% confidence interval on the mean CO2 when x6 = 3200 and x7 = 7.2 is (64.50, 90.57).
## 5.3 90% Prediction Interval for Actual CO2

We calculate the 90% prediction interval for an actual CO2 value when x6 = 3200 and x7 = 7.2.
```{r}
predict(model,interval="prediction",newdata=data.frame(x6=3200,x7=7.2),level=.9)
```
Conclusion: The 90% prediction interval for new CO2 is (56.13, 98.94).

# 6. Refitting the Model Using Only x6

We now refit the model using only x6 as the regressor.
```{r}
model.2 <-lm(y~x6,data=dat)
summary(model.2)

anova(model.2)
```

## 6.1 Testing the Significance of the Reduced Model

We calculate the p-value and rejection region for the reduced model.
```{r}
1-pf(43.766,1,25)
alpha=0.05
qf(1-alpha,1,25)
```
Interpretation:

    F-Test Statistic:
        Fobs=9.74Fobs​=9.74
        Since Fobs>Fcritical=4.24Fobs​>Fcritical​=4.24, we reject the null hypothesis.

    Conclusion:
        The model is effective in explaining the relationship between x6 and y, indicating that x6 is a significant predictor.

# 7. Comparing Models
## 7.1 95% Confidence Intervals for β₆

We compare the 95% confidence intervals for β₆ in both models.
```{r}
confint(model.2,"x6")
confint(model,"x6")
```

# Conclusion:

    For Model 2 (x6 only), the 95% CI for β₆ is (0.0134, 0.0254).
    For Model 1 (x6 and x7), the 95% CI for β₆ is (0.0129, 0.0242).
    The lengths of the confidence intervals are similar, suggesting that the contribution of x7 is not as significant compared to x6 alone.